---
title: "FHEPS Graphics for Final Report"
author: "Patrick Spauster"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: show
---

```{r}
library(tidyverse)
library(tidycensus)
library(tigris)
library(airtabler)
library(sf)
library(DT)
library(urbnthemes)
library(urbnmapr)

set_urbn_defaults(style = "print")

knitr::opts_chunk$set(echo = TRUE)
```

This section adapted from the airtabler documentation: https://github.com/bergant/airtabler

To begin, generate the airtable API key from your Airtable account page: http://airtable.com/account

Airtabler functions will read the API key from environment variable AIRTABLE_API_KEY. While you're at it, if you don't have a census api key, add that to the environment too.  You can sign-up for a key here:https://api.census.gov/data/key_signup.html

To start R session with the initialized environvent variable create an .Renviron file in your home directory by running the following code in the console:

"usethis::edit_r_environ(scope = "project")"

Then add the following text to the .renviron file with the API key you got from the airtable website

"AIRTABLE_API_KEY=your_api_key_here"
"census_api_key=your_api_key_here"

Save the .rEnviron file and restart your R session.

To check where your home is, type path.expand("~") in your R console. Use normalizePath("~/your_file_path") to set it if it's not correct


Pull the airtable data in list form
```{r}
airtable_pull <- airtable(base = "appalJm1WCpSYqwZw", tables = "ALL_clean")

```
get it into a dataframe
```{r}
all_reports <- airtable_pull$ALL_clean$select_all()

```

read in census data
```{r}
nyc_counties <- c("005", "047", "061", "081", "085") #fips for NYC counties

rents_2019 <- get_acs(
  geography = "tract",
  year = 2019,
  state = "NY",
  county = nyc_counties,
  geometry = TRUE,
  variables = "B25064_001E" #med_gross_rent
) %>% 
  mutate(voucher_amount = factor(
            case_when(
                estimate < 1580 ~ "Below voucher amount",
                estimate >= 1580 ~ "Above voucher amount",
                is.na(estimate) ~ "No rent data"), 
            levels = c("Below voucher amount", "Above voucher amount", "No rent data")
  )
  )
```

filter to only cityfheps that have address data
```{r}
FHEPS_reports <- all_reports %>% 
  filter(voucher_type == "CityFHEPS",
         !is.na(long),
         !is.na(lat)) %>% 
  st_as_sf(coords = c("long", "lat")) %>% 
  st_set_crs(st_crs(rents_2019))

```


spatial join
```{r}

FHEPS_tracts <- st_intersection(x = FHEPS_reports, y = rents_2019)
```

Reports by neighborhood median rent
``` {r}
datatable(data = count(as.data.frame(FHEPS_tracts), voucher_amount),
          caption = "Reports in tracts with Median Rents above and below the FHEPS voucher amount",
          colnames = c("Tract Median Rent", "Number of Reports"))

```


Set colors
```{r}
#B6B0A9 light gray
#85837F gray
#9D4350 maroon
#DF9900 orange
#253955 blue
#535840 dark green
#93A081 light green
#E9DBC5 cream
#FECE39 gold
#466CA1 light blue

map_tricolor <- c("#93A081", "#9D4350", "#E9DBC5")
gold <- "#FECE39"
green_ramp <- colorRampPalette(c("#93A081", "#535840"))
green_scale <- greenramp(10)

nyc_water <- nyc_counties %>% 
  map(~area_water("NY", .x)) %>% 
  reduce(rbind) %>% 
  st_set_crs(st_crs(rents_2019))

```

Plot map
```{r}

ggplot()+
  geom_sf(data = rents_2019, mapping = aes(fill = voucher_amount), color = NA)+
  scale_fill_manual(values = map_tricolor) +
  geom_sf(data = nyc_water, mapping = aes(), fill = "white", color = NA)+
  geom_sf(data = FHEPS_reports, mapping = aes(), color = gold, show.legend = TRUE)+
  theme(
    panel.grid = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text = element_blank(),
  )



```

```{r}

ggsave("FHEPS+voucher_amount.png",
       device = "png",
       width = 6.5,
       height = 4.5)

```



Read in opportunity atlas data
```{r}

location <- "C:/Users/patri/Documents/SOI data project" #change this to put the file somewhere on your computer

if (!file.exists(paste0(location, "/tract_outcomes_simple.csv"))) {
  print("downloading data")

  download.file(url = "https://opportunityinsights.org/wp-content/uploads/2018/10/tract_outcomes_simple.csv", paste0(location, "/opportunity_outcomes_simple.csv"))

}

opportunity_data <- read_csv(paste0(location, "/opportunity_outcomes_simple.csv"),
                             col_types = cols(state = col_character(),
                                              county = col_character(),
                                              tract = col_character())) %>% 
  mutate(GEOID = paste0(
    str_pad(state, 2, "0", side = "left"),
    str_pad(county, 3, "0", side = "left"),
    str_pad(tract, 6, "0", side = "left")
  ))


```

```{r}

opportunity_sf <- left_join(rents_2019, opportunity_data, by = "GEOID")

```

```{r}
ggplot()+
  geom_sf(data = filter(opportunity_sf, !is.na(kfr_pooled_pooled_p25*100)), mapping = aes(fill = kfr_pooled_pooled_p25*100), color = NA)+
  scale_fill_gradient(low = "#DCFFCD", high = "#373C2C") +
  geom_sf(data = filter(opportunity_sf, is.na(kfr_pooled_pooled_p25*100)), mapping = aes(), fill = "#E9DBC5", color = NA)+
  geom_sf(data = nyc_water, mapping = aes(), fill = "white", color = NA)+
  geom_sf(data = FHEPS_reports, mapping = aes(), color = gold, show.legend = TRUE)+
  theme(
    panel.grid = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text = element_blank(),
  )+
  guides(fill = guide_colourbar(barwidth = 10, barheight = 0.5))


```

```{r}
ggsave("FHEPS+opportunity.png",
       device = "png",
       width = 6.5,
       height = 4.5)
```

